\systemnamenospace: Head Orientation Based Selection\\in Physical Spaces

### 2014-08-06 doc/sui14/applications.tex
cleared old changes macros, added names, copyright footer, acks to draft for camera ready

% \bjoern{1 is problematic - because it sounds like that the head worn device is in the end not worth it and you should just use your phone, even if it takes longer. I'm not sure we can confidently say something about museums or body shops - this sounds speculative. Needs section needs to be rephrased., e.g. as in: Our solution succeeds at X because of Y. This suggests that future applications in domains like Z could be especially promising.}
\changes{While our system enables users to select and control the example devices successfully, work remains in devising appropriate interfaces for complex devices. Users rated ease of use higher for the lamp and fan which had simple, discrete on/off actions, and lower for the TV control, which had more options. Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}. 
We leave addressing this fundamental usability challenge of wearable devices with near-eye displays to future work.}
%\changes{Though our solution successfully solves the problem of {\em target selection}, we have not fully addressed the issue of controlling smart devices. One user reported that \studyquote{for more complex appliances (TV), interaction was more difficult}. A separate work on interaction techniques for head-worn computing devices is needed to address these issues, while we keep our focus on target selection.} \ben{ Suggest to remove this one: When the device is more complicated, it is possible that a better solution may be a hybrid -- to select the appliance with HOBS and control it with a phone.} \claire{ Is it possible to say that the number of complicated devices that you want to remote control is limited almost exclusively to the tv? I can't think of too many of these devices and our system works very well with more simple devices, which seem to be more common.} \sean{ I wouldn't state that. We could be challenged if the reviewers bring up smart fridge, microwave, etc.}

>>>>

% \bjoern{1 is problematic - because it sounds like that the head worn device is in the end not worth it and you should just use your phone, even if it takes longer. I'm not sure we can confidently say something about museums or body shops - this sounds speculative. Needs section needs to be rephrased., e.g. as in: Our solution succeeds at X because of Y. This suggests that future applications in domains like Z could be especially promising.}
While our system enables users to select and control the example devices successfully, work remains in devising appropriate interfaces for complex devices. Users rated ease of use higher for the lamp and fan which had simple, discrete on/off actions, and lower for the TV control, which had more options. Multiple participants remarked that the difficulty was based on the affordances of Glass: \studyquote{most of the difficulty I had with Glass came from having to navigate the interface on the tiny screen with the touch pad}. 
We leave addressing this fundamental usability challenge of wearable devices with near-eye displays to future work.
%\changes{Though our solution successfully solves the problem of {\em target selection}, we have not fully addressed the issue of controlling smart devices. One user reported that \studyquote{for more complex appliances (TV), interaction was more difficult}. A separate work on interaction techniques for head-worn computing devices is needed to address these issues, while we keep our focus on target selection.} \ben{ Suggest to remove this one: When the device is more complicated, it is possible that a better solution may be a hybrid -- to select the appliance with HOBS and control it with a phone.} \claire{ Is it possible to say that the number of complicated devices that you want to remote control is limited almost exclusively to the tv? I can't think of too many of these devices and our system works very well with more simple devices, which seem to be more common.} \sean{ I wouldn't state that. We could be challenged if the reviewers bring up smart fridge, microwave, etc.}

### 2014-08-06 doc/sui14/interaction.tex
cleared old changes macros, added names, copyright footer, acks to draft for camera ready

%!TEX root = sui14.tex
\section{IR for head orientation-based targeting}
\changes{In this section, we present our hardware platform (IR targeting) and interaction model (head-orientation based) that we will use throughout our iterative designs.}

@@

We hypothesize that infrared (IR) emitters are a good technology match for head orientation selection, since they emit light within a given angle, resulting in a cone in front of the emitter where the light is visible. IR LEDs with many different beam angles are commercially available. \changes{We augment Google Glass with a 940nm 5mm IR emitter with $10^\circ$ beam angle (OSRAM SFH 4545). The emitter is controlled by an additional microcontroller which communicates with Google Glass through Bluetooth radio, since Glass does not directly support hardware modification (see Figure~\ref{fig:glass}). Target devices use Vishay TSOP38238 IR receivers. Data is encoded using standard IR remote protocols at 38.0kHz.}

\changes{IR signals are used for initial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send information such as target IDs and signal strength from targets to back to Glass. We have chosen the commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose (see Figure~\ref{fig:architecture}). This architecture was mostly chosen for reasons of expediency and we do not claim optimality for prototyping decisions. Future head-mounted devices could clearly integrate IR emitters; other wireless techniques (WiFi, Bluetooth) can also be used.}

>>>>

%!TEX root = sui14.tex
\section{IR for head orientation-based targeting}
In this section, we present our hardware platform (IR targeting) and interaction model (head-orientation based) that we will use throughout our iterative designs.

@@

We hypothesize that infrared (IR) emitters are a good technology match for head orientation selection, since they emit light within a given angle, resulting in a cone in front of the emitter where the light is visible. IR LEDs with many different beam angles are commercially available. We augment Google Glass with a 940nm 5mm IR emitter with $10^\circ$ beam angle (OSRAM SFH 4545). The emitter is controlled by an additional microcontroller which communicates with Google Glass through Bluetooth radio, since Glass does not directly support hardware modification (see Figure~\ref{fig:glass}). Target devices use Vishay TSOP38238 IR receivers. Data is encoded using standard IR remote protocols at 38.0kHz.

IR signals are used for initial line-of-sight targeting; subsequently, we use bi-directions wireless communication to send information such as target IDs and signal strength from targets to back to Glass. We have chosen the commercial off-the-shelf ZigBee implementation (XBee based on 802.15.4 radio) for this purpose (see Figure~\ref{fig:architecture}). This architecture was mostly chosen for reasons of expediency and we do not claim optimality for prototyping decisions. Future head-mounted devices could clearly integrate IR emitters; other wireless techniques (WiFi, Bluetooth) can also be used.

### 2014-08-06 doc/sui14/intro.tex
cleared old changes macros, added names, copyright footer, acks to draft for camera ready

\changes{Today, companies like Samsung and Whirlpool are making smart appliances with companion applications that use smartphones as {\em universal remote controls}. With these applications, the user can select a device from a list in order to control it with a device-specific user interface. However, this method faces {\em naming} issues (i.e. ``what do we name the lamp on the left?'') and {\em scaling} issues as the number of controlled devices increases. These solutions also present a necessarily flawed mapping from the positions of the appliances in the rich, 3-dimensional world to their place in a 1D or 2D list presented on the screen. }
%
%% previous approaches are limited
Past research has used direct aiming at target devices in space with phones to overcome these problems~\cite{beigl_point_1999,patel_2-way_2003}. Such techniques have a few drawbacks: the aiming device first has to be retrieved; the user's hands have to be free for operation; and the user's visual attention is split between looking down at a screen and out at targets in the world.

>>>>

Today, companies like Samsung and Whirlpool are making smart appliances with companion applications that use smartphones as {\em universal remote controls}. With these applications, the user can select a device from a list in order to control it with a device-specific user interface. However, this method faces {\em naming} issues (i.e. ``what do we name the lamp on the left?'') and {\em scaling} issues as the number of controlled devices increases. These solutions also present a necessarily flawed mapping from the positions of the appliances in the rich, 3-dimensional world to their place in a 1D or 2D list presented on the screen.
%
%% previous approaches are limited
Past research has used direct aiming at target devices in space with phones to overcome these problems~\cite{beigl_point_1999,patel_2-way_2003}. Such techniques have a few drawbacks: the aiming device first has to be retrieved; the user's hands have to be free for operation; and the user's visual attention is split between looking down at a screen and out at targets in the world.

### 2014-08-06 doc/sui14/iteration1.tex
cleared old changes macros, added names, copyright footer, acks to draft for camera ready

\vfill
\subsection{Technique}
In our first implementation, we use IR for scanning as described in the previous section. For the refinement stage, we simply show a list of the subset of targets that have received IR signals on the Glass display. Users swipe to select from that list and tap to confirm the intended target.

@@

As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided). While both interfaces were judged similarly on overall ease of connecting, \changes{IR selection was perceived to be more direct and pleasurable. One user noted benefit of {\em IR selection}  \studyquote{allowing users to focus on the targeted objects instead of the screen}. \achal{This seems rather weak... Don't really need all the quotes.} One subject called it \studyquote{natural to interact with things just by looking at them}. Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}}.

>>>>

\subsection{Technique}
In our first implementation, we use IR for scanning as described in the previous section. For the refinement stage, we simply show a list of the subset of targets that have received IR signals on the Glass display. Users swipe to select from that list and tap to confirm the intended target.

@@

As for the qualitative feedback, 11 out of 14 participants preferred IR selection over list mode (two preferred list, one was undecided). While both interfaces were judged similarly on overall ease of connecting, IR selection was perceived to be more direct and pleasurable. One user noted benefit of {\em IR selection}  \studyquote{allowing users to focus on the targeted objects instead of the screen}. \achal{This seems rather weak... Don't really need all the quotes.} One subject called it \studyquote{natural to interact with things just by looking at them}. Another mentioned that \studyquote{it's really convenient that what I'm looking at is what I'm targeting}.

### 2014-08-06 doc/sui14/relatedwork.tex
cleared old changes macros, added names, copyright footer, acks to draft for camera ready

%!TEX root = sui14.tex
\vfill
\ben{make sure no section title appears at the end of any column.}

@@

\subsection{Head and Gaze Input}
\changes{Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}. However, it is notable that human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.}
%
%Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}. 
%
Prior work often focused on head orientation for controlling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.

@@

\subsection{Area Cursors}
\changes{
In 2D area cursors for GUIs, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}. We argue that head orientation pointing has analogous characteristics (limited pointing performance and accuracy). Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}. Similar ideas have also been extended into 3D to provide selection with progressive refinement in 3D scenes~\cite{bacim2013design}. All area and space cursors necessitate disambiguation when there are multiple targets and no clear winner. This paper describes the trade-offs between several disambiguation approaches.

%We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, which we call {\em scanning}, users move so the activation area intersects with the target object (and possibly other, unintended targets). In the {\em refinement} phase, they adjust so only the intended target will be selected. Many disambiguation techniques are possible for refinement -- this paper describes the trade-offs between several of them.
}

@@

\subsection{Vision- and Projection-Based Target Selection}
\changes{Many alternative solutions for detecting devices in contained spaces
rely on computer-vision recognition of printed tags on devices. Unfortunately,
these methods either impose significant constraints on the camera used for
detection~\cite{Bokode}, require large or obtrusive tags~\cite{Dataglyphs}, or
are designed to work specifically at short distances~\cite{CyberCode}. Passive
markers also cannot show visual feedback in the environment.  Handheld
projectors can both display a user interface in space and communicate control
information optically, e.g., by encoding information temporally (using Gray
codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG
Lamps~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared
spectrum in SideBySide~\cite{willis_sidebyside:_2011}). Our solution is similar
in spirit but requires only small, low-cost IR emitters and detectors.
\bjoern{need a stronger statement here}} \achal{Tried emphasizing the smaller
requirements, not sure what else we can say.}

>>>>

%!TEX root = sui14.tex
\vfill
%\ben{make sure no section title appears at the end of any column.}

@@

\subsection{Head and Gaze Input}
Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}. However, it is notable that human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.
%
%Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}. 
%
Prior work often focused on head orientation for controlling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.

@@

\subsection{Area Cursors}

In 2D area cursors for GUIs, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}. We argue that head orientation pointing has analogous characteristics (limited pointing performance and accuracy). Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}. Similar ideas have also been extended into 3D to provide selection with progressive refinement in 3D scenes~\cite{bacim2013design}. All area and space cursors necessitate disambiguation when there are multiple targets and no clear winner. This paper describes the trade-offs between several disambiguation approaches.

%We conceptualize area pointing as a two-stage process: in the {\em coarse} phase, which we call {\em scanning}, users move so the activation area intersects with the target object (and possibly other, unintended targets). In the {\em refinement} phase, they adjust so only the intended target will be selected. Many disambiguation techniques are possible for refinement -- this paper describes the trade-offs between several of them.

@@

\subsection{Vision- and Projection-Based Target Selection}
Many alternative solutions for detecting devices in contained spaces
rely on computer-vision recognition of printed tags on devices. Unfortunately,
these methods either impose significant constraints on the camera used for
detection~\cite{Bokode}, require large or obtrusive tags~\cite{Dataglyphs}, or
are designed to work specifically at short distances~\cite{CyberCode}. Passive
markers also cannot show visual feedback in the environment.  Handheld
projectors can both display a user interface in space and communicate control
information optically, e.g., by encoding information temporally (using Gray
codes in Picontrol~\cite{schmidt_picontrol:_2012} and RFIG
Lamps~\cite{raskar_rfig_2004}) or spatially (using QR codes in the infrared
spectrum in SideBySide~\cite{willis_sidebyside:_2011}). Our solution is similar
in spirit but requires only small, low-cost IR emitters and detectors.

### 2014-08-06 doc/sui14/sui14.tex
cleared old changes macros, added names, copyright footer, acks to draft for camera ready

%% EXAMPLE BEGIN -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
% \toappear{Permission to make digital or hard copies of all or part of this work for personal or classroom use is 	granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than ACM must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from permissions@acm.org. \\\textbf{}
% {\emph{CHI'14}}, April 26--May 1, 2014, Toronto, Canada. \\
% Copyright \copyright~2014 ACM ISBN/14/04...\$15.00. \\
% DOI string from ACM form confirmation}
%% EXAMPLE END -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
\toappear{Submitted for review to SUI 2014.}

% Arabic page numbers for submission. 
% Remove this line to eliminate page numbers for the camera ready copy
\pagenumbering{arabic}

>>>>

%% EXAMPLE BEGIN -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
% \toappear{Permission to make digital or hard copies of all or part of this work for personal or classroom use is 	granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than ACM must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from permissions@acm.org. \\\textbf{}
% {\emph{CHI'14}}, April 26--May 1, 2014, Toronto, Canada. \\
% Copyright \copyright~2014 ACM ISBN/14/04...\$15.00. \\
% DOI string from ACM form confirmation}
%% EXAMPLE END -- HOW TO OVERRIDE THE DEFAULT COPYRIGHT STRIP -- (July 22, 2013 - Paul Baumann)
\toappear{Permission to make digital or hard copies of part or all of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for third-party components of this work must be honored. For all other uses, contact the owner/author(s). Copyright is held by the author/owner(s).\\
{\emph{SUI'14}}, October 4–5, 2014, Honolulu, HI, USA.\\
ACM 978-1-4503-2820-3/14/10.\\
\changes{TODO: DOI string goes here.}}

% Arabic page numbers for submission. 
% Remove this line to eliminate page numbers for the camera ready copy
%\pagenumbering{arabic}
